Sky Tomography for 3D Aerosol Distribution Recovery

### 2013-07-23 article.tex
bookmark (returned the theoretical background section into the main article)

Lightfield and integral imaging~\cite{bishop,horstmeyer,kim,Ng1948}
samples the optical radiance distribution in location and
direction. It is mainly used in small-scale setups. This paper deals
with a huge scale, up to Earth's size.  The atmosphere allows light to
pass through in multiple locations and directions. Light is affected
by this medium. Therefore, in this paper we lay out a principle to
recover this 3D medium using measured and modeled lightfields
(Fig.~\ref{fig:front}).
\begin{figure}[t!]
  \centering
  \yoavcomment{\includegraphics[width=\columnwidth]{images/front_img.pdf}}
  \caption{\small The sky is photographed from multiple viewpoints
    using wide angle (hemispherical) cameras, in a very wide area. The
    three dimensional distribution of aerosols in the atmosphere
    affects the images. This distribution is then recovered by a
    dedicated tomography formulation.}
  \label{fig:front}
\end{figure}

Projecting this large object to various directions is similar to
tomography in other scientific domains. However, the formulation here
is distinct. Most tomography setups have a controlled and/or
multidirectional radiation source~\cite{gorbunov,messer}. In contrast,
our source is the uncontrolled, unidirectional Sun. Moreover, typical
tomography relies on a linear model~\cite{gregson}: the pixel value is
a linear combination of components along a line of sight (LOS), or a
multiplicative combination (linearized by a logarithm). Linear
tomography can detect volcanic gas~\cite{wright}, which absorbs
UV. Other gases in the atmosphere absorb IR. However, in visible
light, atmospheric attenuation is typically dominated by {\em
  scattering} by particles (aerosols), rather than absorption by
gasses. Thus, in our problem, we cannot rely on direct illumination,
but on sunlight scattered into the LOS.  The model turns out to be
{\em nonlinear, yet tractable}.

@@

\subfile{theoretic_back}

@@

\section{Image Acquisition Principle}
\label{sec:setup}

The atmosphere has a non-uniform distribution of scattering
aerosols. We seek to recover this 3D volumetric distribution. To solve
this problem, we formulate a process analogous to tomographic
reconstruction. The scene is projected to multiple directions. Thus, a
system captures integrated radiance as a function of direction and 2D
projected location. Any voxel in the atmosphere contributes to the
integrated radiance in multiple directional measurements
(Fig.~\ref{fig:groundgrid}). The measurements then provide a large
number of constraints, to recover the distribution: the density and
type of aerosols, per voxel.
\begin{figure*}[t!]
  \begin{center}
    \includegraphics[width=\linewidth]{images/groundtomog24.pdf}
  \end{center}
  \caption{\small Lightfield imaging through a volumetric distribution
    in the atmosphere, using ground-based cameras. Our tomographic
    reconstruction follows the acquisition. The atmosphere is analyzed
    in a voxel grid.}
  \label{fig:groundgrid}
\end{figure*}
Earth's radiation is sampled both spatially and angularly from
space~\cite{breon,diner,Martonchikc,vanMol}. However, current
multiview spaceborne systems are very coarse. For example, aerosol
retrieval in the most complex instrument, MISR, requires spatial
blocks that are about $17\times 17$km wide, each observed from just 9
views along-track. This practically leads to vertical radiative
transfer analysis, degenerating atmospheric analysis to
1D. Experimental airborne instruments under
development~\cite{baxter,dinerDavis10} may provide better, if still
coarse resolution once they are operational.

@@

\section{Forward Model}
\label{sec:skymodel}

@@

The recovery task is phrased as optimization of an objective function,
that fits the image-formation model (\cref{eq:bigIc,eq:bigIA}) to
measured photographs $\{ {\bm i}^{\rm measured}_c\}_{c=1}^{N_{\rm
    views}}$:
\begin{equation}
  \label{eq:minobjectiveA}
  \hat{\bm n} =
  \argmin_{{\bm n}} F({\bm n})
\end{equation}
where
\begin{equation}
  \label{eq:objectiveA}
  F({\bm n})
  = \sum_{c=1}^{N_{\rm views}}
  \| {\bm i}^{\rm measured}_c - {\bm i}_c({\bm n})\|^2_2
\end{equation}
This is solved using standard optimization tools.  The gradient of $F$
with respect to ${\bm n}$ is
\begin{align}
  \Grad{{\bm n}} F = -2\sum_{c=1}^{N_{\rm views}}
  \transpose{\left[{\bm J}_{{\bm i}_c}({\bm n})\right]} [{\bm i}^{\rm
    measured}_c - {\bm i}_c({\bm n})] \;.
  \label{eq:gradient1}
\end{align}
Here the matrix ${\bm J}_{{\bm i}_c}({\bm n})$ is the Jacobian of the
vector ${\bm i}_c$ with respect to ${\bm n}$. A close-form expression
of our Jacobian is
\begin{align}
  {\bm J}_{{\bm i}_c}({\bm n}) &= L^{\rm TOA}({\bm A}-{\bm B}){\bm C}
  ,
  \label{eq:gradient2}
\end{align}
where
\begin{align}
  \label{eq:gradient3}
  {\bm A} &= \varpi^{\rm aerosol}\sigma^{\rm aerosol}
  \OpDiag{ P_g^{\rm aerosol}(\vect{\Phi}^{\rm scatter}_c)} \nonumber\\
  {\bm B} &= {\sigma}^{\rm aerosol}\transpose{{\bm D}_c} \nonumber\\
  & \;\;\; \OpDiag{[\tilde{\vect{\alpha}}^{\rm air}_c + \varpi^{\rm
      aerosol}\sigma^{\rm aerosol} P^{\rm aerosol}_g(\vect{\Phi}^{\rm
      scatter}_c) \odot{\bm n}
    ]} \nonumber \\
  {\bm C} &= \OpDiag{\exp[-(\vect{\tau}_c^{\rm air} + {\sigma}^{\rm
      aerosol} {\bm D}_c {\bm n})]} \transpose{{\vect{\Pi}}_c}
\end{align}
For a detailed derivation of \cref{eq:gradient2,eq:gradient3}, we
refer the reader to \cref{sec:gradient-derivation}.

>>>>

Lightfield and integral imaging~\cite{bishop,horstmeyer,kim,Ng1948}
samples the optical radiance distribution in location and
direction. It is mainly used in small-scale setups. This paper deals
with a huge scale, up to Earth's size.  The atmosphere allows light to
pass through in multiple locations and directions. Light is affected
by this medium. Therefore, in this paper we lay out a principle to
recover this 3D medium using measured and modeled lightfields
(Fig.~\ref{fig:groundgrid}).
\begin{figure*}[t!]
  \begin{center}
    \includegraphics[width=\linewidth]{images/groundtomog24.pdf}
  \end{center}
  \caption{\small Lightfield imaging through a volumetric distribution
    in the atmosphere, using ground-based cameras. Our tomographic
    reconstruction follows the acquisition. The atmosphere is analyzed
    in a voxel grid.}
  \label{fig:groundgrid}
\end{figure*}
Projecting this large object to various directions is similar to
tomography in other scientific domains. However, the formulation here
is distinct. Most tomography setups have a controlled and/or
multidirectional radiation source~\cite{gorbunov,messer}. In contrast,
our source is the uncontrolled, unidirectional Sun. Moreover, typical
tomography relies on a linear model~\cite{gregson}: the pixel value is
a linear combination of components along a line of sight (LOS), or a
multiplicative combination (linearized by a logarithm). Linear
tomography can detect volcanic gas~\cite{wright}, which absorbs
UV. Other gases in the atmosphere absorb IR. However, in visible
light, atmospheric attenuation is typically dominated by {\em
  scattering} by particles (aerosols), rather than absorption by
gasses. Thus, in our problem, we cannot rely on direct illumination,
but on sunlight scattered into the LOS.  The model turns out to be
{\em nonlinear, yet tractable}.

@@

\section{Related Work}
\label{sec:literature}

{\em Lightfield}, or {\em integral}
imaging~\cite{Ng2006,bishop,Ng1948}, samples the radiance distribution
in location and direction. It has so far been used in small-scale
setups. Our proposed research deals with huge scale, up to Earth's
size. Capturing the lightfield in very large scales requires uncommon
sensors. Sampling Earth both spatially and angularly is done by the
Multiangle Imaging SpectroRadiometer (MISR)~\cite{Martonchikc} from
space (\cref{fig:misr}), the Airborne Multiangle SpectroPolarimetric
Imager (AirMSPI)~\cite{dinerDavis10}, and other spaceborne and
airborne instruments.
\begin{figure}[t]
  \centering
  \includegraphics[width=0.90\textwidth]{misr.pdf}
  \caption{\small{[Left] NASA's MISR instrument, developed by the Jet
      Propulsion Laboratory (JPL), provides multiangular,
      multispectral images of Earth.  It scans Earth from 9 view
      angles, at resolution of $\approx 300$m.  [Right] A series of
      MISR images from different viewing angles showing an assortment
      of clouds over Florida.}  }
  \label{fig:misr}
\end{figure}

Dehazing strives to improve quality of images taken through a
semi-transparent medium, be it murky liquid~\cite{kim,fuchs},
haze~\cite{fattal,Hschechner2}, or fog~\cite{narasimhan2}.  In remote
sensing~\cite{kokhan}, imaging through air is associated with {\em
  atmospheric correction}, based on {\em aerosol retrieval}.  In this
methods the medium is usually assumed to be uniform or at least
plane-parallel.  We, however, seek 3D recovery.  Another major
difference is that these methods see the translucent medium as a
disturbance that algorithms need to overcome, while in our research it
is the medium that we want to recover.

Computed Tomography (CT) refers to the reconstruction of a 3D object
from a collection of projections of the object using any kind of
penetrating radiation. CT finds wide use in medical imaging but is
also valuable in other scientific fields, like
geophysics~\cite{Kazahaya2008}, material science~\cite{Baruchel2000},
astrophysics, and more.  Imaging of gaseous materials is also the
target of extensive CT research~\cite{Price2001,Cosofret2009}. CT is
based on an image formation model that describes the physical process
underlying the projected measurements.  Most image formation models
assume either absorption or emission processes. Other image formation
models assume dominant diffusion by scattering~\cite{Boas2001}.

CT uses several different methods for reconstructing the underlying
medium. Transform algorithms are based on the Fourier and Radon
transforms. Both relate the projection (or its Fourier transform) to
the medium (or its Fourier transform). Solving for the medium involves
calculating the inverse transform in an iterative process.  These
algorithms assume a linear image formation model and require large
amount of projections to give reasonable results. The simultaneous
algebraic reconstruction technique
(SART)~\cite{KakAvinashC.andSlaney2001} is another common method which
is used when the image formation model is too big for explicit matrix
representation. Stochastic Tomography~\cite{gregson} uses a different
approach where the reconstruction is done in the space of projections
using a probabilistic process based on random walk. This bypasses the
need to discretize the signal, usually represented over a grid, and
the use of large matrices to represent the formation model.

Our research topic is closely related to optical
tomography~\cite{Arridge}. Optical tomography uses light to visualize
tissue, usually from close range.  Due to the highly scattering nature
of biological tissue, optical tomography uses diffusive approximation
in the image formation model.  Another method, based on scattering, is
the X-Ray Scatter CT~\cite{Aviles2011} which measures the absorption
and scattering coefficients of the medium separately.

However, the formulation of our problem is {\em different} than prior
situations. Most tomography setups have a controlled and/or
multidirectional radiation source~\cite{messer}. In contrast, our
source is the uncontrolled, unidirectional Sun. Moreover, typical
tomography relies on a linear model: the pixel value is a linear
combination of components along a line of sight (LOS), or a
multiplicative combination which is linearized by a logarithm.
However, photographs of the atmosphere are a mixture of {\em
  scattering}, and absorption. The model turns out to be {\em
  nonlinear, yet tractable}.

@@

\section{Theoretical Background}
\label{sec:theor-backgr}

\noindent {\bf Extinction}: Sun rays (SR) irradiate a small volume
that includes particles of a certain type.  Each particle has an {\em
  extinction cross section} $\sigma$ (units \si{\meter\squared}) for
interacting with the irradiance. The volume includes many particles of
this type. The denser the volume, the stronger is the extinction.  The
density of the particles is $n$ (units
\si[sticky-per]{\per\cubic\metre}). Per unit volume, the {\em
  extinction coefficient} $\beta$ (units \si{\per\meter}) is
\begin{align}
  \beta= \sigma n \;.
  \label{eq:extinctc}
\end{align}
The volume has infinitesimal length $dl$. Then, the relative portion
of extinct SR irradiance is the unitless differential {\em optical
  depth}
\begin{align}
  d\tau= \beta dl=\sigma n dl \;.
  \label{eq:extinct}
\end{align}
The optical depth aggregates in extended propagation:
\begin{align}
  \tau= \int d\tau=\int \beta dl=\int \sigma n dl \;.
  \label{eq:tau}
\end{align}
Through an attenuating volume, the {\em transmittance} exponentially
decays with the optical depth:
\begin{align}
  t=\exp^{-\tau}
  \label{eq:beer-lambert}
\end{align}

\noindent {\bf Scattering}: Interaction of a single particle with the
irradiance is by absorption and scattering. The weight of scattering
(to all directions), relative to the total extinction is given by the
unitless {\em single scattering albedo} $\varpi$ of the particle. The
{\em scattering coefficient} $\alpha$ (units \si{\per\meter}) is
\begin{align}
  \alpha= \varpi\beta=\varpi \sigma n \;.
  \label{eq:alph}
\end{align}
Scattering is generally anisotropic. Its angular distribution is
determined by a {\em phase function} $P$. Part of the light scatters
towards a camera's line of sight (LOS), as illustrated in
\cref{fig:groundgrid}. The angle between the SR and LOS is the
scattering angle $\Phi^{\rm scatter}$. The phase function $P(\Phi^{\rm
  scatter})$ is normalized: its integral over all solid angles is
unit. The {\em angular scattering coefficient} $\tilde\alpha$ (units
\si[sticky-per]{\per\meter\steradian}) is
\begin{align}
  \tilde\alpha(\Phi^{\rm scatter}) = \varpi\beta P(\Phi^{\rm scatter})
  = \varpi \sigma n P(\Phi^{\rm scatter}).
  \label{eq:alphabasic}
\end{align}

The phase function is often approximated by a parametric expression,
as the Henyey-Greenstein~\cite{Cornette1995} function,
\begin{align}
  P_g(\Phi^{\rm scatter})\approx \frac{3} {8\pi} \frac{(1 -
    g^2)(1+\mu^2)} {(2+g^2)(1 + g^2 - 2g\mu)^\frac{3}{2}}
  \label{eq:aerosol_scatter}
\end{align}
where $g$ is an {\em anisotropy parameter} and
\begin{align}
  \mu\equiv \cos \Phi^{\rm scatter}.
  \label{eq:mu}
\end{align}

\noindent {\bf Air molecules}: Scattering by air molecules follows the
{\em Rayleigh} model:
\begin{align}
  P^{\rm air}\left[\mu(k)\right] &\approx \frac{3}{16\pi}(1+\mu^2)
  \label{eq:rayleighP}
\end{align}
and $\varpi^{\rm air}$=1. Air molecular density $n^{\rm air}$ falls
off approximately exponentially with altitude, with a
characteristic~\cite{Levi1980} falloff height $H^\mathrm{air}=8\
\si{\km}$. Consequently~\cite{Levi1980}, the coefficients for
extinction and scattering by air molecules are modeled by
\begin{align}
  \alpha^{\rm air}(h, \lambda)=\beta^{\rm air}(h, \lambda) \approxeq
  \frac{1.09 \times 10^{-3}}{\lambda^4}
  \exp(-h/H^\mathrm{air}) %\left[-\frac{h}{H^\mathrm{air}}\right].
  \label{eq:rayleighbeta}
\end{align}

\section{Methodology}
\label{sec:methodology}

The atmosphere has a non-uniform distribution of scattering
aerosols. We seek to recover this 3D volumetric distribution. To solve
this problem, we formulate a process analogous to tomographic
reconstruction. The scene is projected to multiple directions. Thus, a
system captures integrated radiance as a function of direction and 2D
projected location. Any voxel in the atmosphere contributes to the
integrated radiance in multiple directional measurements
(Fig.~\ref{fig:groundgrid}). The measurements then provide a large
number of constraints, to recover the distribution: the density and
type of aerosols, per voxel.
Earth's radiation is sampled both spatially and angularly from
space~\cite{breon,diner,Martonchikc,vanMol}. However, current
multiview spaceborne systems are very coarse. For example, aerosol
retrieval in the most complex instrument, MISR, requires spatial
blocks that are about $17\times 17$km wide, each observed from just 9
views along-track. This practically leads to vertical radiative
transfer analysis, degenerating atmospheric analysis to
1D. Experimental airborne instruments under
development~\cite{baxter,dinerDavis10} may provide better, if still
coarse resolution once they are operational.

@@

\section{Single-Scattering Model}
\label{sec:single-scatt-model}

@@

The recovery task is phrased as optimization of an objective function,
that fits the image-formation model (\cref{eq:bigIc,eq:bigIA}) to
measured photographs $\{ {\bm i}^{\rm measured}_c\}_{c=1}^{N_{\rm
    views}}$:
\begin{equation}
  \label{eq:minobjectiveA}
  \hat{\bm n} =
  \argmin_{{\bm n}} F({\bm n})
\end{equation}
where
\begin{equation}
  \label{eq:objectiveA}
  F({\bm n})
  = \sum_{c=1}^{N_{\rm views}}
  \| {\bm i}^{\rm measured}_c - {\bm i}_c({\bm n})\|^2_2 + \eta \cdot \Psi({\bm n})
\end{equation}
Here $\Psi({\bm n})$ is a regularization term that expresses some
prior knowledge on the solution and $\eta$ is the regularization
coefficient.  We enforce smoothness of the solution by using the
\emph{Laplacian} operator.
\begin{equation}
  \label{eq:regularizer}
  \Psi({\bm n}) = \| \nabla^2{\bm n}\|^2_2
\end{equation}
We solve~\cref{eq:minobjectiveA} using standard optimization tools.
The gradient of $F$ with respect to ${\bm n}$ is
\begin{align}
  \Grad{{\bm n}} F = -2\sum_{c=1}^{N_{\rm views}}
  \transpose{\left[{\bm J}_{{\bm i}_c}({\bm n})\right]} [{\bm i}^{\rm
    measured}_c - {\bm i}_c({\bm n})] + 2 \cdot \eta \cdot
  \nabla^2\nabla^2{\bm n} \;.
  \label{eq:gradient1}
\end{align}
Here the matrix ${\bm J}_{{\bm i}_c}({\bm n})$ is the Jacobian of the
vector ${\bm i}_c$ with respect to ${\bm n}$, and we make use of the
symmetry of the \emph{Laplacian} operator. A close-form expression of
our Jacobian is
\begin{align}
  {\bm J}_{{\bm i}_c}({\bm n}) &= L^{\rm TOA}({\bm A}-{\bm B}){\bm C}
  ,
  \label{eq:gradient2}
\end{align}
where
\begin{align}
  \label{eq:gradient3}
  {\bm A} &= \varpi^{\rm aerosol}\sigma^{\rm aerosol}
  \OpDiag{ P_g^{\rm aerosol}(\vect{\Phi}^{\rm scatter}_c)} \nonumber\\
  {\bm B} &= {\sigma}^{\rm aerosol}\transpose{{\bm D}_c} \nonumber\\
  & \;\;\; \OpDiag{[\tilde{\vect{\alpha}}^{\rm air}_c + \varpi^{\rm
      aerosol}\sigma^{\rm aerosol} P^{\rm aerosol}_g(\vect{\Phi}^{\rm
      scatter}_c) \odot{\bm n}
    ]} \nonumber \\
  {\bm C} &= \OpDiag{\exp[-(\vect{\tau}_c^{\rm air} + {\sigma}^{\rm
      aerosol} {\bm D}_c {\bm n})]} \transpose{{\vect{\Pi}}_c}
\end{align}
For a detailed derivation of \cref{eq:gradient2,eq:gradient3}, we
refer the reader to \cref{sec:gradient-derivation}.
